---
title: "Prediction Assignment"
author: "gspinnadauhu"
date: "April 27, 2017"
output: html_document
---
#### Data
Loading all packages needed for analysis.
```{r warning=FALSE, verbose=FALSE, message=FALSE}
pkgs<-c("ggplot2","car","caret","lmtest","AppliedPredictiveModeling","plyr","dplyr","ElemStatLearn","forecast","gridExtra","lubridate","magrittr","pander","corrplot","ISLR","parallel","e1071","randomForest")
for (i in 1:length(pkgs)){
        library(pkgs[i],character.only = TRUE)
}
```

Loading training and quiz sets.
```{r warning=FALSE, verbose=FALSE, message=FALSE}
trainURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("data")){
  dir.create("data")
}
download.file(trainURL,destfile="./data/train.csv")
unsplit<-read.csv("./data/train.csv",na.strings=c("","NA"))
download.file(testURL,destfile="./data/test.csv")
quiz<-read.csv("./data/test.csv",na.strings=c("","NA"))
```

The outcome variable is classe, which is divided into 5 levels A through E, defined as:
* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)
Classe A represents the correct way of performing the exercise, the other 4 classes represents common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har


Splitting the training set into a test and train set, as the test set will be used for the quiz portion of the assignment.
```{r}
set.seed(55442)
#creating data partition
inTrain<-createDataPartition(y=unsplit$classe,times=1,p=.7,list=FALSE)
#splitting into train and test sets
train<-unsplit[inTrain,]
test<-unsplit[-inTrain,]
```

The outcome variable is classe, which is divided into 5 levels A through E, defined as:
* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)
Classe A represents the correct way of performing the exercise, the other 4 classes represents common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har

#### Preproccessing
Some basics about the training set.
```{r}
str(train)
```

Going to remove variables that are not likely related to outcome, like user_name, timestamps, etc.
```{r}
train.reduced<-train[,-c(1,2,3,4,5,6,7)]
test.reduced<-test[,-c(1,2,3,4,5,6,7)]
```

Checking the amount of missing data and if the missing data are random or not random.
```{r}
missingpcts<-train.reduced %>% 
   summarise_each(funs(100*mean(is.na(.))))
```

All variables which are missing data are missing the same % of data, which could mean that for most execercises we are not getting any readings from these sensors, so we exclude these variables, since they could otherwise skew results.
```{r}
train.final<-train.reduced[,missingpcts<.8]
test.final<-test.reduced[,missingpcts<.8]
```

Plotting correllation amongst predictors.
```{r}
corrplot(cor(train.final[,-53]),method="color")
```

There appear to be some clusters with highly correllated variables. Some of them seem to be from similar sensors, i.e. dumbell, forearm etc.
Going to use principal components analysis to reduce set of variable to lower number of less correllated variables in train and test set.
```{r}
pca_train<-preProcess(train.final[,-53], method="pca")
train.transformed<-predict(pca_train,newdata=train.final[,-53])
test.trainsformed<-predict(pca_train,newdata=test.final[,-53])
```

#### Model
Since the outcome variable is a 5-level factor, we will use a **Random Forest** prediction model.
```{r}
set.seed(33222)
rand.forest<-train(y=train.final$classe,x=train.transformed,method="rf",ntree=50,tuneLength=1)
rand.forest
```

#### Prediction
Using our rf-model on the test.transformed set and looking at the confusion matrix.
```{r}
pred.rf<-predict(rand.forest,test.trainsformed)
confusionMatrix(test.final$classe,pred.rf)
```

The model accuracy is 97.23% and the out of sample error is therefor 2.77%.

#### Quiz Prediction
First, performing the same pre-processing on the quiz set as on the train/test sets.
```{r}
quiz.reduced<-quiz[,-c(1,2,3,4,5,6,7)]
quiz.final<-quiz.reduced[,missingpcts<.8]
quiz.transformed<-predict(pca_train,newdata=quiz.final)
```
Now, creating the predictions.
```{r}
pred.quiz<-predict(rand.forest,quiz.transformed)
pred.quiz
```