---
title: "Prediction Assignment"
author: "gspinnadauhu"
date: "April 27, 2017"
output: html_document
---
#### Data
Loading training and test sets.
```{r warning=FALSE}
trainURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("data")){
  dir.create("data")
}
download.file(trainURL,destfile="./data/train.csv")
train<-read.csv("./data/train.csv")
download.file(testURL,destfile="./data/test.csv")
test<-read.csv("./data/test.csv")
```

Loading all packages needed for analysis.
```{r warning=FALSE}
pkgs<-c("ggplot2","car","caret","lmtest","AppliedPredictiveModeling","plyr","dplyr","ElemStatLearn","forecast","gridExtra","lubridate","magrittr","pander")
for (i in 1:length(pkgs)){
        library(pkgs[i],character.only = TRUE)
}
```

The outcome variable is classe, which is divided into 5 levels A through E, defined as:
* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)
Classe A represents the correct way of performing the exercise, the other 4 classes represents common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har

#### Exploratory Analysis
Since 'classe' is the outcome, we first look at its properties.

Next, making some plots between index and outcome variable, class.

First 