---
title: "Prediction Assignment"
author: "gspinnadauhu"
date: "April 27, 2017"
output: html_document
---
#### Data
Loading all packages needed for analysis.
```{r warning=FALSE, verbose=FALSE, message=FALSE}
pkgs<-c("ggplot2","car","caret","lmtest","AppliedPredictiveModeling","plyr","dplyr","ElemStatLearn","forecast","gridExtra","lubridate","magrittr","pander","corrplot","ISLR","parallel")
for (i in 1:length(pkgs)){
        library(pkgs[i],character.only = TRUE)
}
```

Loading training and quiz sets.
```{r warning=FALSE, verbose=FALSE, message=FALSE}
trainURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("data")){
  dir.create("data")
}
download.file(trainURL,destfile="./data/train.csv")
unsplit<-read.csv("./data/train.csv",na.strings=c("","NA"))
download.file(testURL,destfile="./data/test.csv")
quiz<-read.csv("./data/test.csv",na.strings=c("","NA"))
```

The outcome variable is classe, which is divided into 5 levels A through E, defined as:
* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)
Classe A represents the correct way of performing the exercise, the other 4 classes represents common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har


Splitting the training set into a test and train set, as the test set will be used for the quiz portion of the assignment.
```{r}
set.seed(55442)
#creating data partition
inTrain<-createDataPartition(y=unsplit$classe,times=1,p=.7,list=FALSE)
#splitting into train and test sets
train<-unsplit[inTrain,]
test<-unsplit[-inTrain,]
```

The outcome variable is classe, which is divided into 5 levels A through E, defined as:
* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)
Classe A represents the correct way of performing the exercise, the other 4 classes represents common mistakes.
Read more: http://groupware.les.inf.puc-rio.br/har

#### Exploratory Analysis
Some basics about the training set.
```{r}
str(train)
```

Going to remove variables that are not likely related to outcome, like user_name, timestamps, etc.
```{r}
train.reduced<-train[,-c(1,2,3,4,5,6,7)]
test.reduced<-test[,-c(1,2,3,4,5,6,7)]
```

Checking the amount of missing data and if the missing data are random or not random.
```{r}
missingpcts<-train.reduced %>% 
   summarise_each(funs(100*mean(is.na(.))))
```

All variables which are missing data are missing the same % of data, which could mean that for most execercises we are not getting any readings from these sensors, so we exclude these variables, since they could otherwise skew results.
```{r}
train.final<-train.reduced[,missingpcts<.8]
test.final<-test.reduced[,missingpcts<.8]
```


Plotting predictors.
```{r}
featurePlot(x=train.final,y=train.final$classe,plot="pairs")
```

